A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-05-13 paper/main.tex
Updates from Overleaf

% defining affiliations
\newcommand{\affuofa}{University of Arizona, 933 N. Cherry Ave,
    Tucson, AZ 85721, USA}
\newcommand{\affuofu}{Department of Physics and Astronomy, University of Utah, 115 South 1400 East, Salt Lake City, Utah 84112, USA}

@@

@@

\begin{abstract}
    % Statement
    The merger timescales of isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) on cosmologically motivated orbits have not yet been studied in detail, though isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively.
    % While the merger timescales of isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) in cosmological environments have been studied extensively, isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) have not yet been studied in detail.
    It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, nor if they evolve similarly with redshift. 
    % Our solution
    We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time, and explore how different selection criteria impact the mass and redshift dependence of merger timescales. 
    % What are the timescales? 
    We find that the merger timescales for low- and high-mass pairs are the same from $z=0-6$ in the absence of any separation criteria.
    We then present a physically-motivated framework for selecting pairs via a scaled separation criteria, wherein pair separations are scaled by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 \Rvir$).  
    Applying these scaled separation criteria recovers equivalent merger timescales for both mass scales at all redshifts.
    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts leads to a difference in merger rates of up to $\sim1\Gyr$ between low- and high-mass pairs, particularly for $\rsep<150\kpc$. 
    % Importance
    As a result, applying the same merger timescales to physical separation-selected pairs will lead to a bias that systematically over-predicts low-mass galaxy merger rates. 
    % could replace the above line with something softer suggesting the application of scaled selection criteria for studies of merger properties of pairs at different mass scales and redshifts, or something like the below:    
    % Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system. 
    % if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.

@@

\section{Methodology} \label{sec:methods}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009} and the merger tree catalogs generated by the \sublink{} algorithm~\citep{RG2015} from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

Our sample consists of major low-mass and high-mass pairs (stellar mass ratio $1/4 \leq M_{*2}/M_{*1}\leq 1$) that are isolated, but physically associated, as in \citet{Chamberlain2024}.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.

@@

At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass $\MG$
% GB added M_G
\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass. This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Bryan1998} and $\rho_c$ is the critical density of the universe at the time calculated. The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.} 
\begin{align*}
        \mbox{\textbf{low mass:}}&\,\rm \MG = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
        \mbox{\textbf{high mass:}}&\, \rm \MG =1\times 10^{12}- 6.5\times10^{12}\,\Msun.
\end{align*}
Requiring pairs to belong to the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.

@@

\subsection{Orbits} 
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
    \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary as a function of lookback time. 
    (Top) Orbits of pairs that merge before $z=0$.
    (Middle) Orbits of pairs that do not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
    (Bottom) Orbits of pairs that do not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
    Solid lines represent the post-infall orbits, i.e. after the pair share a common FoF group.
    Dashed lines show the pre-infall portion of the orbit. 
    Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. 
    }
    \label{fig:example-orbits}
    \end{center}
\end{figure}

@@

\section{Pair Sample Properties}\label{sec:pairprops}
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{The number of unique low-mass (green) and high-mass (pink) pairs in our orbit catalog as a function of redshift. Orbits are only defined between first infall and the merger redshift (for merging pairs) or $z=0$ (for non-merging pairs). 
    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers), while dashed lines show the number of pairs at a given redshift that will merge before $z=0$.
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{"total"}
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
    \caption{The fraction of low-mass (green) and high-mass (pink) orbits that merge before z=0 shown as a function of redshift. 
    The horizontal line shows a merger fraction of 0.9, or 90\%. 
    High-mass pairs have merger fractions greater than 0.9 for $z>1.1$, and low-mass pairs for $z>1.6$.
    %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
    The sharp decline of the merger fraction to zero at low redshift is a non-physical feature of the simulation ending at $z=0$.
    Only orbits with sufficiently short merger timescales will merge at redshifts $z<\sim1$.}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Pairs}\label{sec:pairprops-num}
The total number of pairs at a given redshift is equal to the number of orbits at that redshift, including both merger and non-merger pairs. 
%We calculate the total number of pairs by selecting all pairs at a given redshift, including merger and non-merger pairs.
A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.

@@

\begin{figure*}[htb!]
    \begin{center}
    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/8_2Dhist.png}
    \caption{The distribution of merger timescales for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
    The colorbars show the percentage of the total number of pairs at each redshift (given in the bottom right corner of each panel) that are in a given separation bin.
    In each of the first panels, the grey horizontal dashed line shows the time remaining in the simulation, above which there are no mergers. 
    %
    We find:
    i.) a positive correlation between separation and the time until merger at each selected redshift -- the lowest separation pairs tend to have the shortest merger timescales; 
    ii.) given a redshift and separation, the low-mass merger timescale is longer than the high-mass merger timescale, i.e. low-mass pairs at $z=1$ with separations $\rsep<100\,\kpc$ merge within $6\, \Gyr$, while high-mass pairs in the same separation range merge within $3\, \Gyr$; 
    iii.) merger timescales are longer for lower redshift pairs; 
    iv.) the spread of merger timescales and separations increase with decreasing redshift. }
    \label{fig:2dhist}
    \end{center}
\end{figure*}

@@

    We find that the merger timescale for low-mass and high-mass pairs is positively correlated with the separation of the pair at a given redshift. 
    The merger timescale also increases with decreasing redshift for a given separation.
    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$. 
    For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass at each redshift. 
    % However, the slope of this correlation decreases at higher redshifts.

@@

@@

        \begin{figure}[tb]
            \centering
            \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale.png}
            \caption{The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs in our merger sample. 
            Shaded regions represent the 1st and 3rd quartile spread on the median. 
            % 
            The thick dotted black line at low redshifts shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the merger timescale that a merging pair can have. 
            The thin black line shows the Hubble time as a function of redshift multiplied by a constant $\beta=0.35$, which approximates the median time until merger from $z=6$ to $z\sim1$.
            %
            The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
            \label{fig:timescales}
        \end{figure}

@@

        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by $\beta=0.35$. 
        We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the specific values of the merger timescale. 
        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=\sim1-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation, from~\citet{Planck2018}. Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}

@@

        The median merger timescale for low-mass pairs does not change significantly for any separation cuts above $150\,\kpc$, which is larger than the average virial radius of low-mass systems at all redshifts. 
        On the other hand, high-mass merger timescales tend to increase as separation increases.
        % This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
        High-mass pairs tend to have higher separations than low-mass pairs, as shown in Sec.~\ref{sec:results-timevsep}, due to the larger size of the high-mass subhalos. 
    \begin{figure*}[htb]
        \centering
        \includegraphics[width=\textwidth]{plots/bet-on-it/8_timescale_sepcut.png}
        \caption{(Top) The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $50,70,100,150,200,\mbox{and }300\,\kpc$ from left to right. 
        (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5\,\Rvir$ from left to right. 
        % 
        The thin black line is the same as shown in Fig.~\ref{fig:timescales}, and goes as the Hubble time $1/H(z)$ multiplied by 0.35 (see Sec.~\ref{sec:results-timevredshift} for additional details).
        The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$. 
        }
        \label{fig:timescales-sep}
    \end{figure*}  
    \subsubsection{Scaled Separation Selected Pairs}
    \label{sec:results-scal}
        We calculate the merger timescales for six additional subsamples of pairs, applying separation criteria that scale with both redshift and mass.
        The scaled separation, which we define in Section~\ref{sec:methods}, is the separation of a pair divided by the virial radius of the pair's FoF group, $\Rvir$. 
        Scaled separation criteria select the equivalent \textit{fraction} of the full volume of each FoF group, regardless of mass or redshift, whereas physical separation criteria will select the same volume around each primary, with no account for the growth of dark matter halos or spatial distributions of satellites. 
        The scaled separation selected pairs have separations greater than $10\,\kpc$ and less than $[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$. 
        As in the previous section, the selection criteria are applied at each redshift independently.

@@

            In recent years, some work has aimed to standardize pair selection criteria for future studies, in particular, to facilitate a fair comparison between theory and observations. 
            One such study is~\citet{Ventou2019}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of (both physical and projected) separation and relative velocity. 
            They find that selecting pairs with projected separations between $5\leq r_{\rm proj} \leq 50\,\kpc$ and projected relative velocities of $v_{\rm proj}\leq300\,\kms$ selects all pairs that have greater than 30\% chance of merging before $z=0$. 
            These selection criteria are calibrated to select a population of pairs that are fairly likely to merge.
            Using their projected separation vs velocity analysis of merger probabilities, they also develop a weighting scheme to determine the likelihood of merger as a function of projected separation and velocity, which provides a way to more accurately calibrate merger fractions from close pair fractions. 
            While their study considers the impact of stellar mass and redshift on the probability of merger, their suggested pair selection criteria do not vary with mass or redshift, which we have shown is crucial for interpretting and comparing pair fractions and merger timescales.

@@

            % \textit{RG2015} 
            % the reason I want to talk about the RG15 paper is because its an example of trying to put galaxy merger rates on the same plots, and using the same selection criteria at high and low masses may have also resulted in such a discrepancy, IN ADDITION to the problems that low mass morphology studies face when trying to use morphological signatures for low-mass merger fraction studies.

@@

            These results imply that the timescales used to convert observational merger fractions (determined either by weighted close pair fractions or by morphologically selected pairs) are sensitive to the mass ranges and redshifts at which they are valid, and attempting to make comparisons between theoretical predictions and observational measurements can lead to seemingly incompatible results. 
            % Since our observational samples themselves will be able to also self-consistently apply merger timescales to close pair fractions using our suggested scaled separation criteria.
            Our work provides a unique framework that will allow for more robust comparisons of theoretical and observational samples.

@@

            We provide a step by step example of this application as follows. First, compute the associated dark matter masses from the observed stellar masses of the pair using a SMHM relation~\citep[in this study, we employ that of][]{Moster2013}.
            % In particular, given an observation pair with stellar masses $\ms1$ and $\ms2$, we can use the SMHM relation to calculate the associated dark matter masses, $M_{h1}$ and $M_{h2}$, of the galaxy pair. 
            Next, calculate the virial radius of a dark matter halo with a virial mass equal to the sum of the dark matter masses of the pair.\footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy.}
            Finally, determine the physical separation for each pair individually that corresponds to the scaled separation criteria adopted in the study.

            The process of computing the dark matter halo masses of the observed galaxies typically introduces systematic error to an observational study.
            However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos. Furthermore, the SMHM relationship of~\cite{Moster2013} was constructed to account for observational stellar mass errors.
            When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled in a similar fashion as performed in~\chambe{} to derive the spread associated with the abundance matching process.

            As a concluding note, we remind readers that our results are explicitly for isolated pairs. 
            One strength of this approach is that we have been able to study dynamics that are inherent to the pairs themselves, rather than those that are a product of the environment. 
            Low-mass and high-mass pair merger timescales in high density environments may evolve differently than in our findings. 
            We leave the extension of this analysis to more high-density environments as the focus of future work.

\section{Summary and Conclusions}\label{sec:conclusions}
    % what we did and how
    In this study, we construct a sample of the orbits of low-mass ($10^8 <\ms{}<5\times10^{9}\Msun$) and high-mass ($5\times10^{9}<\ms{}<10^{11}\Msun$) major pairs (stellar mass ratio $> 1:4$) from $z=0-6$ in the TNG100 simulation.
    Orbits of pairs, i.e. the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo first share a common FoF group (i.e., `first infall') to either $z=0$ or the redshift at which the pair merges.

>>>>

% defining affiliations
\newcommand{\affuofa}{University of Arizona, 933 N. Cherry Ave,
    Tucson, AZ 85721, USA}
\newcommand{\affuofu}{Department of Physics and Astronomy, University of Utah, 115 South 1400 East, Salt Lake City, Utah 84112, USA}
\newcommand{\affuva}{Department of Astronomy, University of Virginia, 530 McCormick Road, Charlottesville, VA 22904, USA}

@@

\author[0000-0002-5653-0786]{Paul Torrey}
\affiliation{\affuva}

@@

\begin{abstract}
    % Statement
    The merger timescales of isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) on cosmologically motivated orbits have not yet been studied in detail, though isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively.
    % While the merger timescales of isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) in cosmological environments have been studied extensively, isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) have not yet been studied in detail.
    It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, 
    %nor
    % GB edited
    or
    if they evolve similarly with redshift. 
    % Our solution
    We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time, and explore how different selection criteria impact the mass and redshift dependence of merger timescales. 
    % What are the timescales? 
    We find that the merger timescales for low- and high-mass pairs are the same from $z=0-6$ in the absence of any separation criteria.
    We then present a physically-motivated framework for selecting pairs via a scaled separation criteria, wherein pair separations are scaled by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 \Rvir$).  
    Applying these scaled separation criteria recovers equivalent merger timescales for both mass scales at all redshifts.
    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts leads to a difference in merger rates of up to $\sim1\Gyr$ between low- and high-mass pairs, particularly for $\rsep<150\kpc$. 
    % Importance
    As a result, applying the same merger timescales to physical separation-selected pairs will lead to a bias that systematically over-predicts low-mass galaxy merger rates. 
    % could replace the above line with something softer suggesting the application of scaled selection criteria for studies of merger properties of pairs at different mass scales and redshifts, or something like the below:    
    % Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system. 
    % if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.

@@

\section{Methodology} \label{sec:methods}
The IllustrisTNG simulation suite~\citep{TNG1,TNG2,TNG3,TNG4,TNG5} is a set of large volume dark-matter-only and full magnetohydrodynamical cosmological simulations consistent with the \textit{Planck 2015} \lcdm{} cosmology~\citep{Planck2015}.

Following~\chambe{}, we use TNG100-~1 (hereafter TNG100), which is the highest resolution full physics run of the ($100\Mpc)^3$ volume.
In particular, we utilize the group catalogs produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009} and the merger tree catalogs generated by the \sublink{} algorithm~\citep{RG2015}.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

Our sample consists of major low-mass and high-mass pairs (stellar mass ratio $1/4 \leq M_{*2}/M_{*1}\leq 1$) that are isolated, but physically associated, as in \citet{Chamberlain2024}.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales from $z=0-6$.

@@

At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass $\MG$
% GB added M_G
\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass. This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Bryan1998} and $\rho_c$ is the critical density of the universe at the time calculated. The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.} 
\begin{align*}
        \mbox{\textbf{low mass:}}&\,\rm \MG = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
        \mbox{\textbf{high mass:}}&\, \rm \MG =1\times 10^{12}- 6.5\times10^{12}\,\Msun.
\end{align*}
Requiring pairs to belong to the same FoF group ensures that the pairs are distant from other massive nearby systems that could perturb the dynamical state of the pair.  
In addition, for each pair, we calculated the Hill Radius of every FoF group with a higher mass within $10Mpc$ of the pair, as a proxy to discern whether the pair lies outside the gravitational sphere of influence of these more massive groups.
We found that over 99\% of our pairs at $z=0$ are more distant than two times the Hill Radius of each FoF group.

@@

\subsection{Orbits} 
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
    % [Example Orbits for Galaxy Pairs in TNG100 Selected at $z=1.5$]
    \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary as a function of lookback time. 
    (Top) Orbits of pairs that merge before $z=0$.
    (Middle) Orbits of pairs that do not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
    (Bottom) Orbits of pairs that do not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
    Solid lines represent the post-infall orbits, i.e. after the pair share a common FoF group.
    Dashed lines show the pre-infall portion of the orbit. 
    Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. 
    }
    \label{fig:example-orbits}
    \end{center}
\end{figure}

@@

\section{Pair Sample Properties}\label{sec:pairprops}
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    % [Number of Unique Pairs in Orbit Catalog]
    \caption{The number of unique low-mass (green) and high-mass (pink) pairs in our orbit catalog as a function of redshift. Orbits are only defined between first infall and the merger redshift (for merging pairs) or $z=0$ (for non-merging pairs). 
    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers), while dashed lines show the number of pairs at a given redshift that will merge before $z=0$.
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
    %[Merger Fraction of Pairs as a Function of Redshift]
    \caption{The fraction of low-mass (green) and high-mass (pink) orbits that merge before z=0 shown as a function of redshift. 
    The horizontal line shows a merger fraction of 0.9, or 90\%. 
    High-mass pairs have merger fractions greater than 0.9 for $z>1.1$, and low-mass pairs for $z>1.6$.
    %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
    The sharp decline of the merger fraction to zero at low redshift is a non-physical feature of the simulation ending at $z=0$.
    Only orbits with sufficiently short merger timescales will merge at redshifts $z<\sim1$.}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Pairs}\label{sec:pairprops-num}
The total number of pairs at a given redshift is equal to the number of orbits at that redshift, including both merger and non-merger pairs. 
A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.

@@

\begin{figure*}[htb!]
    \begin{center}
    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/10_2Dhist_fits.png}
    %[Distribution of Time Until Merger as a Function of Separation at Redshifts $z=0.5,1,2,3,4,5,6$]
    \caption{The distribution of merger timescales for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
    The colorbars show the percentage of the total number of pairs at each redshift (given in the bottom right corner of each panel) that are in a given separation bin.
    In each of the first panels, the grey horizontal dashed line shows the time remaining in the simulation, above which there are no mergers. 
    %
    We find:
    i.) a positive correlation between separation and the time until merger at each selected redshift -- the lowest separation pairs tend to have the shortest merger timescales; 
    ii.) given a redshift and separation, the low-mass merger timescale is longer than the high-mass merger timescale, i.e. low-mass pairs at $z=1$ with separations $\rsep<100\,\kpc$ merge within $6\, \Gyr$, while high-mass pairs in the same separation range merge within $3\, \Gyr$; 
    iii.) merger timescales are longer for lower redshift pairs; 
    iv.) the spread of merger timescales and separations increase with decreasing redshift. }
    \label{fig:2dhist}
    \end{center}
\end{figure*}

@@

    We find that the merger timescale for low-mass and high-mass pairs is positively correlated with the separation of the pair at a given redshift. 
    The merger timescale also increases with decreasing redshift for a given separation.
    For example, pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$. 
    For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass at each redshift. 
    % However, the slope of this correlation decreases at higher redshifts.

@@

    \input{paper/table-slopes}

@@

        \begin{figure}[tb]
            \centering
            \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale.png}
            % [Median Merger Timescale of Low-mass and High-mass Pairs as a Function of Redshift]
            \caption{The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs in our merger sample. 
            Shaded regions represent the 1st and 3rd quartile spread on the median. 
            % 
            The thick dotted black line at low redshifts shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the merger timescale that a merging pair can have. 
            The thin black line shows the Hubble time as a function of redshift multiplied by a constant $\beta=0.35$, which approximates the median time until merger from $z=6$ to $z\sim1$.
            %
            The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
            \label{fig:timescales}
        \end{figure}

@@

        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by $\beta=0.35$. 
        We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the specific values of the merger timescale. 
        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=\sim1-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation, from~\citet{Planck2015}. Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}

@@

        The median merger timescale for low-mass pairs does not change significantly for any separation cuts above $150\,\kpc$, which is larger than the average virial radius of low-mass systems at all redshifts. 
        On the other hand, high-mass merger timescales tend to increase as separation increases.
        % This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
        High-mass pairs tend to have higher separations than low-mass pairs, as shown in Sec.~\ref{sec:results-timevsep}, due to the larger size of the high-mass subhalos. 
    \begin{figure*}[htb]
        \centering
        \includegraphics[width=\textwidth]{plots/bet-on-it/8_timescale_sepcut.png}
        % [Median Merger Timescale as a Function of Redshift for Pairs Selected by Physical and Scaled Separation Criteria]
        \caption{(Top) The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $50,70,100,150,200,\mbox{and }300\,\kpc$ from left to right. 
        (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5\,\Rvir$ from left to right. 
        % 
        The thin black line is the same as shown in Fig.~\ref{fig:timescales}, and goes as the Hubble time $1/H(z)$ multiplied by 0.35 (see Sec.~\ref{sec:results-timevredshift} for additional details).
        The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$. 
        }
        \label{fig:timescales-sep}
    \end{figure*}  
    \subsubsection{Scaled Separation Selected Pairs}
    \label{sec:results-scal}
        We calculate the merger timescales for six additional subsamples of pairs, applying separation criteria that scale with both redshift and mass.
        The scaled separation, which we define in Section~\ref{sec:methods}, is the separation of a pair divided by the virial radius of the pair's FoF group, $\Rvir$. 
        Scaled separation criteria select the equivalent \textit{fraction} of the full volume of each FoF group, regardless of mass or redshift, whereas physical separation criteria will select the same volume around each primary, with no account for the growth of dark matter halos or spatial distributions of satellites. 
        The scaled separation selected pairs have separations greater than $10\,\kpc$ and less than $[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$. 
        As in the previous section, the selection criteria are applied at each redshift independently.

@@

            In recent years, some work has aimed to standardize pair selection criteria for future studies, in particular, to facilitate a fair comparison between theory and observations. 
            One such study is~\citet{Ventou2019}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of (both physical and projected) separation and relative velocity. 
            They find that selecting pairs with projected separations between $5\leq r_{\rm proj} \leq 50\,\kpc$ and projected relative velocities of $v_{\rm proj}\leq300\,\kms$ selects all pairs that have greater than 30\% chance of merging before $z=0$. 
            These selection criteria are calibrated to select a population of pairs that are fairly likely to merge.
            Using their projected separation vs velocity analysis of merger probabilities, they also develop a weighting scheme to determine the likelihood of merger as a function of projected separation and velocity, which provides a way to more accurately calibrate merger fractions from close pair fractions. 
            While their study considers the impact of stellar mass and redshift on the probability of merger, their suggested pair selection criteria do not vary with mass or redshift, which we have shown is crucial for interpreting and comparing pair fractions and merger timescales.

@@

@@

            These results imply that the timescales used to convert observational merger fractions (determined either by weighted close pair fractions or by morphologically selected pairs) are sensitive to the mass ranges and redshifts at which they are valid, and attempting to make comparisons between theoretical predictions and observational measurements can lead to seemingly incompatible results. 
            Our work provides a unique framework that will allow for more robust comparisons of theoretical and observational samples.

@@

            We provide a step by step example of this application as follows. First, compute the associated dark matter masses from the observed stellar masses of the pair using a SMHM relation~\citep[in this study, we employ that of][]{Moster2013}.
            % Paul's comment:
            % you could make this more straight forward, if you want.  Specifically, there is a M_* to R_vir correction that exists in TNG, and you could parameterize and "give" this to the observers. 
            Next, calculate the virial radius of a dark matter halo with a virial mass equal to the sum of the dark matter masses of the pair.
            \footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy. The virial radius can be calculated from
            \begin{equation}\label{eq}
                \Rvir = \sqrt[3]{\frac{3\rm M_{vir}}{4\pi \Delta_{c}\rho_{c}}},
            \end{equation} 
            where $\rho_c$ is the critical density of the Universe, and $\Delta_{c}$ is the overdensity constant~\citep[see][]{Binney2008}.}
            Finally, determine the physical separation for each pair individually that corresponds to the scaled separation criteria adopted in the study.

            The process of computing the dark matter halo masses of the observed galaxies typically introduces systematic error to an observational study.
            When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled, i.e. by computing many realizations of the dark matter mass of each galaxy, in a similar fashion as performed in~\chambe{}, to derive the associated spread introduced by the abundance matching process.

            % However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos. Furthermore, the SMHM relationship of~\cite{Moster2013} was constructed to account for observational stellar mass errors.
            % When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled in a similar fashion as performed in~\chambe{} to derive the spread associated with the abundance matching process.

            As a concluding note, we remind readers that our results are explicitly for isolated pairs, as outlined in Sec.~\ref{sec:methods}. 
            One strength of this approach is that we have been able to study dynamics that are inherent to the pairs themselves, rather than those that are a product of the environment. 
            Low-mass and high-mass pair merger timescales in high density environments may evolve differently than in our findings. 
            We leave the extension of this analysis to more high-density environments as the focus of future work.
